---
title: "CL_analysis"
author: "Meishan"
date: '2023-05-03'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(psych)
library(nlme)
library(fpc)
```
# ####this is an example code of performing clustering on lifestyle behaviors: exercise, cognitive activity, diet, social####
# regress out covariates for your variables (optional, many papers also do not regress out)
```{r}
#get the name of variables
varlist<- names(Lifestyle)
# this renaming is easier for plotting
new_varlist <- c("exercise engagement","cognitive activity","social network","healthy diet adherence")
#compute the residuals
models <- lapply(varlist, function(x) {
  lm(substitute(i ~ age + sex + APOE4_status + Education_years, list(i = as.name(x))), data = df)
})
list_resid = lapply(models, resid)  
#combine multiple lists into table
df_resid = do.call(cbind, list_resid)
df_resid <- as_tibble(df_resid)
#change the variable names
curnames <-names(df_resid)
LS_resid <- df_resid %>% rename_at(vars(curnames), ~ new_varlist)
# Now you should have a table with residual scores across all lifestyle variables

```

# running k-means based on residual scores
```{r}
# run across multiple sizes of k
set.seed(123)
kclusts <- 
  tibble(k = 1:9) %>%
  mutate(
    kclust = map(k, ~kmeans(LS_resid, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, LS_resid)
  )

kclusts
```

# summary outcomes
```{r}
# tidy output across k sizes
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))
```

#scree plot for within-cluster sum of squares, if you need to report it
```{r}

ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + scale_x_continuous(name="K",limits = c(1,9), breaks = c(1:9))+theme(axis.text.x=element_text(size=15), axis.title.x = element_text(size=20), axis.text.y = element_text(size = 15), axis.title.y = element_text(size=20)) 

```

#choose a final k based on the scree plot
```{r}
#I decided cluster=2 based on the plot
set.seed(758)
final_clust2 <- kmeans(LS_resid, centers = 2)
# now you have the final cluster labels for each participants
final_clust2
# #### below  is to plot lifestyle scores by cluster groups
#1 combine lifestyle scores and clustering results
sd <- cbind(LS_resid,cluster=final_clust2$cluster) %>% aggregate(.~cluster, data=., FUN = "sd") %>% pivot_longer(!cluster, names_to = "scale", values_to = "sd")
#2 create df of center of clusters values
centers <- as_tibble(final_clust2[["centers"]])
centers$cluster = c(1,2)

#reshape the data for plotting
centers_long <- centers %>%
  pivot_longer(!cluster, names_to = "scale", values_to = "score")

centers_long$cluster <- as.factor(centers_long$cluster)
centers_long$sd <- sd$sd
# plot out
ggplot(centers_long) +
  geom_bar( aes(x=scale, y=score, fill=cluster), stat="identity", alpha=0.5, position = position_dodge()) +
  geom_errorbar(aes(scale, ymin=score-sd, ymax=score+sd), width=.2,
                 position=position_dodge(.9))

```

#bootstrap k-means to evaluate the stability of your clusters
```{r}
# swap k for each cluster number to evaluate for each number of k
clustBoot <- clusterboot(LS_resid, B = 100,
                         clustermethod = kmeansCBI,
                         k = 2, seed = 12345,
                         showplots = FALSE)
clustBoot # gives jaccard index (if higher then higher stability it is)
```